Online Balanced Graph Partitioning: Optimizations for Large Clusters

### 2020-05-06 paper/disc20clustering.tex
restructuring the proof

\begin{theorem} \label{thm:k=2}
	The algorithm, for $\lambda=\alpha$, is (strictly) 6-competitive.
	\maciek{for what lambda? \mahmoud{done}}
\end{theorem}

@@

Assume on request $\sigma_{t}=\set{u,v}$, 
i.e., at (time) $t$,
Alg collocates the pair $\set{u,v}$.
This means,
the value of $C_{\set{u,v}}$ at $t$,
denoted by $C^t_{\set{u,v}}$, 
reaches $\lambda$ immediately before Alg resets the counter.
I.e.,
$ C^t_{\set{u,v}} = \lambda$.
%	In case the first collocation of $\set{u,v}$ occurs at $t$,
%	if the pair is collocated in the initial configuration then
%	we set $t''=0$,
%	else,
%	we set $ t'' = -\infty, t'=0$.
We denote the set of all requests to a pair $\set{x,y}$ that arrive
during $(t_1,t_2]$ by $\sigma_{\set{x,y}}(t_1,t_2]$,
and its overall cost to OPT by
$\mathit{Cost}_{\mathit{OPT}} (\sigma_{\{u,v\}}(t',t]) $.
%
Observe that for requests in
$R_t := \sigma_{\{u,v'\}}(t'',t'] \cup \sigma_{\{u,v\}}(t',t]$ 
and the entire duration of the input sequence (i.e., $(0,m]$),
OPT must pay a total cost
%We denote the cost w.r.t~$r_t$ by
$\mathit{Cost}_{\mathit{OPT}} (R_t) \geq 
	c_{R_t} := \min{ \{ 2\lambda, \lambda+\alpha, 2\alpha \} }$,
for serving (some of) these requests remotely or (and) collocating (one of) the pairs,
regardless of its choices.
Then the cost per pair, in particular the cost w.r.t.~$\{u,v\}$, is
$\mathit{Cost}_{\mathit{OPT}} (\sigma_{\{u,v\}}(t',t]) \geq c_{R_t} / 2$.
%
However,
we can charge this cost to OPT only at $t=t_0$,
i.e.~the first time Alg collocates $\{u,v\}$,
and not at any consequent time $\tilde{t} > t_0$ that Alg collocates $\{u,v\}$.
%	i.e.,
%	when Alg collocates the two pairs a second time.
That is,
we charge OPT a cost incurred by requests in
$\sigma_{\{u,v'\}}(t'',t']$ or $\sigma_{\{u,v\}}(t',t]$,
if and only if  OPT incurs the cost during  intervals $(t'', t']$ or $(t', t]$,
respectively.	
Latter,
this ensures us that we do not charge OPT for a cost more than twice.
%%
Based on this observation,
in the proof that follows,
we distinguish three cases of OPT's choices
while it serves requests in	$R_t $.

\todo{what is causing this vertical space before proof block??}

\begin{proof}[Proof of Theorem \ref{thm:k=2}]
	If $t$ is the first time that Alg collocates $\{u,v\}$ then  OPT pays
	$\lambda$ for serving requests in $\sigma_{\set{u,v}}[0,t]$ (remotely),
	or $\alpha$ for collocating the pair and
	serving (some of) the requests with cost 0.
	Therefore in this case,
	OPT pays
	$\mathit{Cost}_{\mathit{OPT}} (\sigma_{\{u,v\}}(0,t]) \geq  \min{\{ \lambda,\alpha \}}$.
	%
	Else, this is not the first time and
	Alg collocates $\{u,v\}$ also some time prior to $t$.
	Let $t' < t$ be the  time when Alg separates $\set{u,v}$
	in order to collocate some (intersecting) pair
	$\{x,y\} \neq \{u,v\}, \{x,y\} \cap \{u,v\} \neq \emptyset$, 
	e.g.,
	$\{x,y\}=\{u,v'\}$.
	%	I.e., $ C^{t'}_{\set{u,v'}} = \lambda$.
	Similarly,
	if $t'$ is not the first time that Alg  collocates $\{u,v'\}$ 
	then let $t''<t'$ be the latest time before $t'$ when Alg separates $\set{u,v'}$.
	Else,
	$\{u,v'\}$ is not collocated in the initial configuration and we let $t''=0$.
	By these definitions,
	we have
	$| \sigma_{\set{u,v'}}(t'',t'] | = | \sigma_{\set{u,v}}(t',t] |  =  \lambda$.
	%
	The overall cost of requests in $\sigma_{\set{u,v}}(t',t]$,
	and the moving cost,
	incurred by Alg is
	$\lambda + 2\alpha$.
	OPT pays,
	$\mathit{Cost}_{\mathit{OPT}} (R_t) := \mathit{Cost}_{\mathit{OPT}} (\sigma_{\{u,v'\}}(t'',t']) 
	+ \mathit{Cost}_{\mathit{OPT}} (\sigma_{\{u,v\}}(t',t]) $.
	We distinguish three (non-exhaustive) cases for the cost to OPT.
	Latter,
	 we argue that OPT cannot save cost by deviating from these cases.
	\begin{enumerate}[label=\roman*.]
		\item
		OPT collocates $\{u,v'\}$ at  $t''$, $\{u,v\}$  at $t'$,
		and keeps them so during $(t'',t']$ and $(t',t]$, respectively.
		Therefore it serves all requests in $R_t$ with no additional cost and
		$ c_1 (R_t) := 2\alpha$.
		\item
		OPT has $\{u,v\}$ ($\{u,v'\}$) already collocated at $t'$ ($t''$) as well as during $(t',t]$ ($(t'',t]$),
		and has $\{u,v'\}$ ($\{u,v\}$) separated during $(t'',t']$ ($(t',t]$).
		Hence,
		it pays only $\lambda$ for serving all requests in $R_t$ and
		$c_2 (R_t) := \lambda$
		\item
		OPT has $\{u,v'\}$ ($\{u,v\}$) already collocated at $t''$ ($t'$)  and also during $(t'',t']$ ($(t', t])$.
		Then it pays 0 for $\sigma_{\{u,v'\}}[t'',t']$ ($(\sigma_{\{u,v\}}[t',t]$), and
		$c_3 (R_t) := \alpha$.
	\end{enumerate}
	If OPT's choice is one of these cases,
	then
	$\mathit{Cost}_{\mathit{OPT}} (R_t) \geq
		\min{\{ c_1(R_t), c_2(R_t), c_3(R_t) \}}  =
		\min{ \{ \lambda, \alpha \}}$.
	Otherwise,
	OPT deviates from these cases,
	e.g.~by collocating a pair after paying for some of requests to this pair.
	However,
	no deviation can decrease the cost to OPT.
	Therefore in any case,
	$\mathit{Cost}_{\mathit{OPT}} (R_t) \geq  \min{ \{ \lambda, \alpha \}}$.

	It remains to bound the cost incurred by Alg and OPT for requests that do not lead to collocations until the  end of the sequence at $t=m$.
	Assume $\{u,v\}$ is not collocated at $t=m$
	and $0 < C^{m}_{ \{u,v\} } < \lambda $,
	which means Alg pays $C^{m}_{ \{u,v\} }$
	for requests to $\{u,v\}$ so far unaccounted.
	If $\{u,v\}$ is not collocated in the initial configuration
	and Alg never collocates it,
	then $ C^{m}_{ \{u,v\} } =| \sigma_{\{u,v\}}[1,m] |$.
	OPT pays
	$\mathit{Cost}_{\mathit{OPT}} (\sigma_{\{u,v\}}[1,t]) 
	\geq \min{ \{ \alpha, C^{m}_{ \{u,v\} } \} }$,
	for collocating this pair or (and) paying for (some of) requests in $\sigma_{\{u,v\}}[1,t]$.
	Else,
	either $\{u,v\}$ is collocated in the initial configuration
	or Alg collocates it at some point.
	Then there exists an intersecting pair $\set{u,v'}$
	that is collocated by Alg at $t' < m$ (separating $\{u,v\}$).
	We define times $t'' < t' < m$ similarly to the case where Alg collocates $\{u,v\}$.
	Let $R_{\{u,v\}} := \sigma_{\{u,v'\}} (t'',t'] \cup \sigma_{\{u,v\}} (t',m]$.
	Then,
	we apply cases (i) to (iii) in an analogous way,
	replacing $\lambda$ with $C^{m}_{ \{u,v\}}$ in the analysis.
	Thus, 
	we obtain
	$\mathit{Cost}_{\mathit{OPT}} (R_{\{u,v\}}) 
	\geq  \min{ \{ C^{m}_{ \{u,v\}}, \alpha \}}$.
	The overall cost to Alg is

@@

	Next, we sum up all costs incurred by OPT.
	Observe that by definition,
	$R_{t'} \cap R_t = \sigma_{\{u,v'\}}[t'',t']$.
	This means,
	the cost of r.h.s.~to OPT is counted at most twice in each of  the expressions
	$\mathit{Cost}_{\mathit{OPT}} (R_{t'}) + \mathit{Cost}_{\mathit{OPT}} (R_t)$
	and  
	$\mathit{Cost}_{\mathit{OPT}} (R_{t'}) + \mathit{Cost}_{\mathit{OPT}} (R_{\{u,v\}})$.
	Hence,
	for all collocations performed by Alg,
	and for final costs at $t=m$,
	OPT pays at least 
	$\frac{1}{2}(
	\sum_{ t \in [1,m] } \mathit{Cost}_{\mathit{OPT}} (R_t) +
	\sum_{\{u,v\}} \mathit{Cost}_{\mathit{OPT}} (R_{\{u,v\}})
	) $.
	Let
	$T := \{ t \in [1,m] ~\vert~ \exists \{u,v\}: C^{t}_{\{u,v\}} = \lambda \}$.
	Then,
	the total cost to OPT is

@@

Thus, we have

>>>>

\begin{theorem} \label{thm:k=2}
	The algorithm, for $\lambda=\alpha$, is (strictly) 6-competitive.
\end{theorem}

@@

Assume on request $\sigma_{t}=\set{u,v}$, 
i.e., at (time) $t$,
Alg collocates the pair $\set{u,v}$.
This means,
the value of $C_{\set{u,v}}$ at $t$,
denoted by $C^t_{\set{u,v}}$, 
reaches $\lambda$ immediately before Alg resets the counter.
I.e.,
$ C^t_{\set{u,v}} = \lambda$.
%	In case the first collocation of $\set{u,v}$ occurs at $t$,
%	if the pair is collocated in the initial configuration then
%	we set $t''=0$,
%	else,
%	we set $ t'' = -\infty, t'=0$.
We denote the set of all requests to a pair $\set{x,y}$ that arrive
during $[t_1,t_2]$ by $\sigma_{\set{x,y}}[t_1,t_2]$,
and its overall cost to OPT by
$\mathit{Cost}_{\mathit{OPT}} (\sigma_{\{u,v\}}[t_1,t_2])$.
%
%Else, this is not the first time and
%Alg collocates $\{u,v\}$ also some time prior to $t$.

If $t$ is not the first time that Alg collocates $\{u,v\}$ then
let $0 < t' < t$ be the latest time when Alg separates $\set{u,v}$
in order to collocate some (intersecting) pair
$\{x,y\} \neq \{u,v\}, \{x,y\} \cap \{u,v\} \neq \emptyset$, 
e.g.,
$\{x,y\}=\{u,v'\}$.
Else,
$t$ is the first time that Alg collocates $\{u,v\}$ and let $t' := 0$.
Similarly,
if $t' > 0$ is not the first time that Alg  collocates $\{u,v'\}$ 
then let $0 < t'' < t'$ be the latest time before $t'$ when Alg separates $\set{u,v'}$.
Else,
$t'$ is the first time that Alg collocates $\{u,v'\}$ and we let $t''=0$.
%

Observe that for requests in
$R_t := \sigma_{\{u,v'\}}(t'',t'] \cup \sigma_{\{u,v\}}(t',t]$ 
and the entire duration of the input sequence (i.e., $(0,m]$),
OPT must pay a total cost
%We denote the cost w.r.t~$r_t$ by
$\mathit{Cost}_{\mathit{OPT}} (R_t) \geq 
	c_{R_t} := \min{ \{ 2\lambda, \lambda+\alpha, 2\alpha \} }$,
for serving (some of) these requests remotely or (and) collocating (one of) the pairs,
regardless of its choices.
Then the cost per pair, in particular the cost w.r.t.~$\{u,v\}$, is
$\mathit{Cost}_{\mathit{OPT}} (\sigma_{\{u,v\}}(t',t]) \geq c_{R_t} / 2$.
%
However,
we can charge this cost to OPT only at $t=t_0$,
i.e.~the first time Alg collocates $\{u,v\}$,
and not at any consequent time $\tilde{t} > t_0$ that Alg collocates $\{u,v\}$.
%	i.e.,
%	when Alg collocates the two pairs a second time.
That is,
we charge OPT a cost incurred by requests in
$\sigma_{\{u,v'\}}(t'',t']$ or $\sigma_{\{u,v\}}(t',t]$,
if and only if  OPT incurs the cost during  intervals $(t'', t']$ or $(t', t]$,
respectively.	
Latter,
this ensures us that we do not charge OPT for a cost more than twice.
%%
Based on this observation,
in the proof that follows,
we distinguish three cases of OPT's choices
while it serves requests in	$R_t $.

\begin{proof}[Proof of Theorem \ref{thm:k=2}]
	First,
	we bound the costs incurred by Alg for requests that
	lead to the collocation of $\{u,v\}$ at time $t \in T$, where
	$T := \{ i \in [1,m] ~\vert~ \exists \{x,y\}: C^{i}_{\{x,y\}} = \lambda \}$.
	By definitions of $t$ and $t'$,
	the overall cost of requests in $\sigma_{\set{u,v}}(t',t]$ incurred by Alg,
	i.e., the total cost of remote serving
	and the moving cost, is
	$\lambda + 2\alpha$.	
	Next,
	we bound the cost incurred by Alg
	for requests that do not lead to collocations until the  end of the sequence at $t=m$.
	Assume $\{u,v\}$ is not collocated at $t=m$
	and $0 < C^{m}_{ \{u,v\} } < \lambda $,
	which means Alg pays $C^{m}_{ \{u,v\} }$
	for these requests $\sigma_{\set{u,v}}(t',m]$.
    Then the overall cost to Alg is

@@

	Next,
	we bound  costs incurred by OPT	for requests that trigger collocation of $\{u,v\}$ at $t \in T$.
	If $t$ is the first time that Alg collocates $\{u,v\}$ then  OPT pays
	$\lambda$ for serving requests in $\sigma_{\set{u,v}}[0,t]$ (remotely),
	or $\alpha$ for collocating the pair and
	serving (some of) the requests with cost 0.
	Therefore in this case,
	OPT pays
	$\mathit{Cost}_{\mathit{OPT}} (\sigma_{\{u,v\}}(0,t]) \geq  \min{\{ \lambda,\alpha \}}$.
	%
	By these definitions,
	we have
	$| \sigma_{\set{u,v'}}(t'',t'] | = | \sigma_{\set{u,v}}(t',t] |  =  \lambda$.
	For requests $R_t := \sigma_{\set{u,v'}}(t'',t']  \cup \sigma_{\set{u,v}}(t',t]$,
	OPT pays
	$\mathit{Cost}_{\mathit{OPT}} (R_t) := \mathit{Cost}_{\mathit{OPT}} (\sigma_{\{u,v'\}}(t'',t']) 
	+ \mathit{Cost}_{\mathit{OPT}} (\sigma_{\{u,v\}}(t',t]) $.
	We distinguish three (non-exhaustive) cases for the cost to OPT.
	Latter,
	 we argue that OPT cannot save cost by deviating from these cases.
	\begin{enumerate}[label=\roman*.]
		\item
		OPT collocates $\{u,v'\}$ at  $t''$, $\{u,v\}$  at $t'$,
		and keeps them so during $(t'',t']$ and $(t',t]$, respectively.
		Therefore it serves all requests in $R_t$ with no additional cost and
		$ c_1 (R_t) := 2\alpha$.
		\item
		OPT has $\{u,v\}$ ($\{u,v'\}$) already collocated at $t'$ ($t''$) as well as during $(t',t]$ ($(t'',t]$),
		and has $\{u,v'\}$ ($\{u,v\}$) separated during $(t'',t']$ ($(t',t]$).
		Hence,
		it pays only $\lambda$ for serving all requests in $R_t$ and
		$c_2 (R_t) := \lambda$
		\item
		OPT has $\{u,v'\}$ ($\{u,v\}$) already collocated at $t''$ ($t'$)  and also during $(t'',t']$ ($(t', t])$.
		Then it pays 0 for $\sigma_{\{u,v'\}}[t'',t']$ ($(\sigma_{\{u,v\}}[t',t]$), and
		$c_3 (R_t) := \alpha$.
	\end{enumerate}
	If OPT's choice is one of these cases,
	then
	$\mathit{Cost}_{\mathit{OPT}} (R_t) \geq
		\min{\{ c_1(R_t), c_2(R_t), c_3(R_t) \}}  =
		\min{ \{ \lambda, \alpha \}}$.
	Otherwise,
	OPT deviates from these cases,
	e.g.~by collocating a pair after paying for some of requests to this pair.
	However,
	no deviation can decrease the cost to OPT.
	Therefore in any case,
	$\mathit{Cost}_{\mathit{OPT}} (R_t) \geq  \min{ \{ \lambda, \alpha \}}$.

	It remains to bound the cost  incurred by OPT for requests to $\{u,v\}$ that do not lead to its collocation until the end of the sequence at $t=m$.
	If $\{u,v\}$ is not collocated in the initial configuration
	and Alg never collocates it,
	then $ C^{m}_{ \{u,v\} } =| \sigma_{\{u,v\}}[1,m] |$.
	OPT pays
	$\mathit{Cost}_{\mathit{OPT}} (\sigma_{\{u,v\}}[1,m]) 
	\geq \min{ \{ \alpha, C^{m}_{ \{u,v\} } \} }$,
	for collocating this pair or (and) paying for (some of) requests in $\sigma_{\{u,v\}}[1,m]$.
	Else,
	either $\{u,v\}$ is collocated in the initial configuration
	or Alg collocates it at some point.
	Then there exists an intersecting pair $\set{u,v'}$
	that is collocated by Alg at $t' < m$ (separating $\{u,v\}$).
	We define times $t'' < t' < m$ similarly to the case where Alg collocates $\{u,v\}$.
	Let $R_{\{u,v\}} := \sigma_{\{u,v'\}} (t'',t'] \cup \sigma_{\{u,v\}} (t',m]$.
	Then,
	we apply cases (i) to (iii) in an analogous way,
	replacing $\lambda$ with $C^{m}_{ \{u,v\}}$ in the analysis.
	Thus, 
	we obtain
	$\mathit{Cost}_{\mathit{OPT}} (R_{\{u,v\}}) 
	\geq  \min{ \{ C^{m}_{ \{u,v\}}, \alpha \}}$.

	Next, we sum up all costs incurred by OPT.
	Observe that by definition,
	$R_{t'} \cap R_t = \sigma_{\{u,v'\}}[t'',t']$.
	This means,
	the cost of r.h.s.~to OPT is counted at most twice in each of  the expressions
	$\mathit{Cost}_{\mathit{OPT}} (R_{t'}) + \mathit{Cost}_{\mathit{OPT}} (R_t)$
	and  
	$\mathit{Cost}_{\mathit{OPT}} (R_{t'}) + \mathit{Cost}_{\mathit{OPT}} (R_{\{u,v\}})$.
	Hence,
	for all collocations performed by Alg,
	and for final costs at $t=m$,
	OPT pays at least 
	$\frac{1}{2}(
	\sum_{ t \in [1,m] } \mathit{Cost}_{\mathit{OPT}} (R_t) +
	\sum_{\{u,v\}} \mathit{Cost}_{\mathit{OPT}} (R_{\{u,v\}})
	) $.
	Then,
	the total cost to OPT is

@@

Finally, we have
